---
title: "ARIMA Demo for MC"
output: html_notebook
---

```{r}
library(forecast)
library(tidyverse)
library(dplyr)

sales <- read_csv("./data/read_summarized_sales_clean_new.csv")
gdp_sectors <- read_csv("./data/read_gdp_sector_pivot_clean.csv")

```


```{r}

sales <- sales %>% filter(`Year-Qtr` != '2024-Q4')

```

```{r}
qtr_sales <- sales |> 
  group_by(`Year-Qtr`) |> 
  summarize(
    Total_Rev = sum(`Total Rev`, na.rm = TRUE), 
    .groups = "drop"
  ) 

sales_gdp <- gdp_sectors |> 
  inner_join(qtr_sales, by = "Year-Qtr")

```



```{r}

cutoff_year = 2022

#Use data before the cutoff year to fit the model

model_tib_gdp <- sales_gdp |> 
  arrange(Year) |> 
  filter(Year < cutoff_year)
```





```{r}
# model_tib_gdp <- sales_gdp
```

We need to convert the variables to time series objects.
```{r}
total_rev_ts <- ts(model_tib_gdp$Total_Rev, start = min(model_tib_gdp$Year), frequency = 4)
gdp_total_ts <- ts(model_tib_gdp$GDP_Total, start = min(model_tib_gdp$Year), frequency = 4)

#ts(model_tib_gdp$Total_Rev,start = min(model_tib_gdp$Year) frequency = 4, start = c(2021, 1))
```

Then, we can use the auto.arima function from the forecast library to fit an arima model.
For the US, it uses one autoregressive term. This means that it is using the prior year's life expectancy as a predictor for this year's value.
```{r}
gdp_sales_model <- auto.arima(
  total_rev_ts, 
  xreg = log(gdp_total_ts)
)

summary(gdp_sales_model)
```

If we want to test for statistical significance, we can use the coeftest function from the lmtest library.
```{r}
library(lmtest)

coeftest(gdp_sales_model)
```

Let's see how well it forecasts for future years. Because of the big dropoff in life expectancy for 2020, we'll only look at years before 2020.
```{r}
# future_GDP <- sales_gdp |> 
#   pull(GDP_Total)
# forecasted_values <- forecast(gdp_sales_model, 
#                               xreg = log(future_GDP)
#                               )



future_GDP <- sales_gdp |> 
  arrange(Year) |> 
  filter(Year >= cutoff_year) |>
  filter(Year < 2024) |> 
  pull(GDP_Total)
forecasted_values <- forecast(gdp_sales_model, 
                              xreg = log(future_GDP)
                              )
         

start_year <- 2022

years <- seq(start_year, start_year + length(future_GDP) / 4 - 1)

quarters <- rep(c("Q1", "Q2", "Q3", "Q4"), length.out = length(future_GDP))

Year_Qtr <- paste(rep(years, each = 4), quarters, sep = "-")
         
         
print(future_GDP)
print(Year_Qtr)
         

```


Let's plot the actual (solid line) vs. the forecasted (dashed line). This also includes the 80% and 95% confidence intervals for the forecasted values.
```{r}
forecast_tib_qtr <- tibble(
  Year_Qtr,
  Total_Rev_Exp_Pred = forecasted_values$mean,
  Lower_80 = forecasted_values$lower[,1],
  Upper_80 = forecasted_values$upper[,1],
  Lower_95 = forecasted_values$lower[,2],
  Upper_95 = forecasted_values$upper[,2]
)




Year_Qtr_Float <- rep(years, each = 4) + match(quarters, c("Q1", "Q2", "Q3", "Q4")) / 4

forecast_tib_qtr$`Year-Qtr-Float` <- Year_Qtr_Float


forecast_tib_qtr <- forecast_tib_qtr |> 
  mutate(`Year-Qtr-Float` = Year_Qtr_Float)

forecast_tib_qtr


```

```{r}


forecast_tib_qtr

```

```{r}
ggplot() +
  geom_line(data = sales_gdp |> 
            filter(Year < 2024),
            aes(x = `Year-Qtr-Float`, y = Total_Rev)
            ) +
  geom_line(data = forecast_tib_qtr, 
            aes(x = Year_Qtr_Float, y = Total_Rev_Exp_Pred ), 
            linetype = "dashed") +
  geom_ribbon(data = forecast_tib_qtr, 
              aes(x = Year_Qtr_Float, ymin = Lower_95, ymax = Upper_95), 
              alpha = 0.5,
              fill = "gray80") +
  geom_ribbon(data = forecast_tib_qtr, 
              aes(x = Year_Qtr_Float, ymin = Lower_80, ymax = Upper_80),
              alpha = 0.5,
              fill = "gray70")

```





















